Supplemental Information B: Full Annotated Analysis and Reproducible Code

### 2021-10-25 Methods-and-Results.Rmd
AOT-P methods added

The full surveys are available in our Online Supplementary Material A and the full data description is available in Online Supplementary Materials B.

@@

# divide into trait and state
dtWorkerSupp$workerWithinBetween <-
  MlTraitState(
    data = dtWorkerSupp$workerInteractionType,
    id = "PID",
    selection =
      c(
        "keyMotiveFulfilled",
        "competence.daily.all",
        "autonomy.daily.all",
        "relatedness.daily.all",
        "thermometerDutch_1",
        "keymotive_fulfillemt_1",
        "competence_1",
        "autonomy_1",
        "relatedness_1"
      )
  )

workerMlCor <-
  MlCorMat(
    data = dtWorkerSupp$workerInteractionType,
    id = "PID",
    selection = c("keyMotiveFulfilled", "competence.daily.all", "autonomy.daily.all", "relatedness.daily.all", "thermometerDutch_1") ,
    labels = c("Core Need", "Competence", "Autonomy", "Relatedness", "Attitudes NL")
  )

workerMlCor %>%
  kable(
    .,
    caption = "Worker: Multilevel Core Variable Descriptives",
    format = "latex",
    booktabs = TRUE,
    align = c("l", rep("c", ncol(.) - 1))
  ) %>%
  footnote(
    general = c(
      "Upper triangle: Between-person correlations;",
      "Lower triangle: Within-person coorelations;",
      "*** p < .001, ** p < .01,  * p < .05"
    )
  ) %>%
  save_kable("Tables/workerVarDescr.tex")

workerOutMlCor <-
  MlCorMat(
    data = dtWorkerSupp$workerInteractionType %>% filter(OutgroupInteraction == "Yes"),
    id = "PID",
    selection = c("keymotive_fulfillemt_1", "competence_1", "autonomy_1", "relatedness_1", "thermometerDutch_1") ,
    labels = c("Core Need", "Competence", "Autonomy", "Relatedness", "Attitudes NL")
  )

@@

\paragraph{Psychological Needs} Irrespective of whether participants had an interaction with Dutch people or not, everyone answered a short series of questions on psychological need fulfillment. However, whereas participants with interactions reported on the need fulfillment during the interaction, people without interactions with Dutch people judged the past daytime period in general. To assess the fulfillment of psychological needs, we included two types of need measurement: (1) the core situational need and (2) general self-determination theory needs.

For the core situational need, we asked participants in an open ended text field: “\textit{What was your most important goal [during the interaction / this morning / this afternoon]?}”. We then used the participant’s open text entry of their core need to see how much this was fulfilled during the interaction or the past daytime period: “\textit{[The interaction / you] fulfilled your goal: [-previous text entry-]}” on a continuous slider scale ranging from strongly disagree (-50) to strongly agree (+50).

We, additionally, included a common measure of three self-determination theory needs \citep[taken from][]{Downie2008}. The items were introduced either by “\textit{During the interaction:}” or “\textit{This morning [/afternoon]:}” and measured autonomy (“\textit{I was myself.}”), competence (“\textit{I felt competent.}”), and relatedness (without intergroup contact “\textit{I had a strong need to belong}”; with intergroup contact: “\textit{I shared information about myself.}” and “\textit{The other(s) shared information about themselves.}”). All items were rated on a continuous slider scale from very little (-50) to a great deal (+50).

\paragraph{Perceived Interaction Quality}

\paragraph{Outgroup Attitudes}

>>>>

The full surveys are available in our Online Supplementary Material A and the full data description is available in Online Supplementary Materials B. Correlations and descriptive statistics of the included variables are available in Table \ref{tab:workerVarDescr} and Table \ref{tab:workerOutVarDescr}.

@@

# divide into trait and state
dtWorkerSupp$workerWithinBetween <-
  MlTraitState(
    data = dtWorkerSupp$workerInteractionType,
    id = "PID",
    selection =
      c(
        "keyMotiveFulfilled",
        "competence.daily.all",
        "autonomy.daily.all",
        "relatedness.daily.all",
        "thermometerDutch_1",
        "keymotive_fulfillemt_1",
        "competence_1",
        "autonomy_1",
        "relatedness_1"
      )
  )
```

```{r workerVarDescr, include=FALSE}
workerMlCor <-
  MlCorMat(
    data = dtWorkerSupp$workerInteractionType,
    id = "PID",
    selection = c("keyMotiveFulfilled", "competence.daily.all", "autonomy.daily.all", "relatedness.daily.all", "thermometerDutch_1") ,
    labels = c("Core Need", "Competence", "Autonomy", "Relatedness", "Attitudes NL")
  )

workerMlCor %>%
  kable(
    .,
    caption = "Worker: Multilevel Core Variable Descriptives",
    format = "latex",
    booktabs = TRUE,
    align = c("l", rep("c", ncol(.) - 1))
  ) %>%
  footnote(
    general = c(
      "Upper triangle: Between-person correlations;",
      "Lower triangle: Within-person coorelations;",
      "*** p < .001, ** p < .01,  * p < .05"
    )
  ) %>%
  save_kable("Tables/workerVarDescr.tex")
```

```{r workerOutVarDescr, include=FALSE}
workerOutMlCor <-
  MlCorMat(
    data = dtWorkerSupp$workerInteractionType %>% filter(OutgroupInteraction == "Yes"),
    id = "PID",
    selection = c("keymotive_fulfillemt_1", "competence_1", "autonomy_1", "relatedness_1", "thermometerDutch_1") ,
    labels = c("Core Need", "Competence", "Autonomy", "Relatedness", "Attitudes NL")
  )

@@

\paragraph{Psychological Needs}Irrespective of whether participants had an interaction with Dutch people or not, everyone answered a short series of questions on psychological need fulfillment. However, whereas participants with interactions reported on the need fulfillment during the interaction, people without interactions with Dutch people judged the past daytime period in general. To assess the fulfillment of psychological needs, we included two types of need measurement: (1) the core situational need and (2) general self-determination theory needs.

For the core situational need, we asked participants in an open ended text field: “\textit{What was your most important goal [during the interaction / this morning / this afternoon]?}”. Then, with reference to the text entry, we asked how much this core need was fulfilled during the interaction or the past daytime period: “\textit{[The interaction / you] fulfilled your goal: [-previous text entry-]}” on a continuous slider scale ranging from strongly disagree (-50) to strongly agree (+50).

We, additionally, included a common measure of three self-determination theory needs \citep[see][]{Downie2008}. The items were introduced either by “\textit{During the interaction:}” or “\textit{This morning [/afternoon]:}” and measured autonomy (“\textit{I was myself.}”), competence (“\textit{I felt competent.}”), and relatedness (without intergroup contact “\textit{I had a strong need to belong}”; with intergroup contact: “\textit{I shared information about myself.}” and “\textit{The other(s) shared information about themselves.}”). All items were rated on a continuous slider scale from very little (-50) to a great deal (+50).

\paragraph{Perceived Interaction Quality}As an explanatory mechanism, we assessed ratings of the perceived interaction quality. As our main measurement, participants rated the statement “\textit{Overall the interaction was …}” on two continuous slider scales measuring pleasantness \citep[from unpleasant (-50) to pleasant (+50)) and meaningfulness (from superficial (-50) to meaningful (+50); both items adapted from][]{Downie2008}.

\paragraph{Outgroup Attitudes}At the end of every daily diary measurement we asked all participants about their current attitudes towards the Dutch – our main dependent variable. To assess the momentary outgroup evaluation we used the common feeling thermometer: “How favorable do you feel towards the Dutch?” \citep[][]{Lavrakas2008}. Participants then rated their attitude on a continuous slider scale from “very cold – 0” through “no feeling – 50” to “very warm – 100”. Both the question phrasing as well as the tick labels were consistent with large-scale panel surveys \citep[e.g.,][]{DeBell2010}.

### 2021-10-25 Methods-and-Results.tex
AOT-P methods added

For the core situational need, we asked participants in an open ended
text field:
``\textit{What was your most important goal [during the interaction / this morning / this afternoon]?}''.
We then used the participant's open text entry of their core need to see
how much this was fulfilled during the interaction or the past daytime
period:
``\textit{[The interaction / you] fulfilled your goal: [-previous text entry-]}''
on a continuous slider scale ranging from strongly disagree (-50) to
strongly agree (+50).

We, additionally, included a common measure of three self-determination
theory needs \citep[taken from][]{Downie2008}. The items were introduced
either by ``\textit{During the interaction:}'' or
``\textit{This morning [/afternoon]:}'' and measured autonomy
(``\textit{I was myself.}''), competence
(``\textit{I felt competent.}''), and relatedness (without intergroup
contact ``\textit{I had a strong need to belong}''; with intergroup
contact: ``\textit{I shared information about myself.}'' and
``\textit{The other(s) shared information about themselves.}''). All
items were rated on a continuous slider scale from very little (-50) to
a great deal (+50).

@@

@@

>>>>

For the core situational need, we asked participants in an open ended
text field:
``\textit{What was your most important goal [during the interaction / this morning / this afternoon]?}''.
Then, with reference to the text entry, we asked how much this core need
was fulfilled during the interaction or the past daytime period:
``\textit{[The interaction / you] fulfilled your goal: [-previous text entry-]}''
on a continuous slider scale ranging from strongly disagree (-50) to
strongly agree (+50).

We, additionally, included a common measure of three self-determination
theory needs \citep[see][]{Downie2008}. The items were introduced either
by ``\textit{During the interaction:}'' or
``\textit{This morning [/afternoon]:}'' and measured autonomy
(``\textit{I was myself.}''), competence
(``\textit{I felt competent.}''), and relatedness (without intergroup
contact ``\textit{I had a strong need to belong}''; with intergroup
contact: ``\textit{I shared information about myself.}'' and
``\textit{The other(s) shared information about themselves.}''). All
items were rated on a continuous slider scale from very little (-50) to
a great deal (+50).

@@

As an explanatory mechanism, we assessed ratings of the perceived
interaction quality. As our main measurement, participants rated the
statement ``\textit{Overall the interaction was …}'' on two continuous
slider scales measuring pleasantness
\citep[from unpleasant (-50) to pleasant (+50)) and meaningfulness (from superficial (-50) to meaningful (+50); both items adapted from][]{Downie2008}.

@@

At the end of every daily diary measurement we asked all participants
about their current attitudes towards the Dutch -- our main dependent
variable. To assess the momentary outgroup evaluation we used the common
feeling thermometer: ``How favorable do you feel towards the Dutch?''
\citep[][]{Lavrakas2008}. Participants then rated their attitude on a
continuous slider scale from ``very cold -- 0'' through ``no feeling --
50'' to ``very warm -- 100''. Both the question phrasing as well as the
tick labels were consistent with large-scale panel surveys
\citep[e.g.,][]{DeBell2010}.
